\documentclass{article}
\usepackage{handout}
\usepackage{algorithm}
\usepackage{myhyperlinks}
\usepackage{graphicx}

%\newcommand{\cmt}[1]{\textsf{[#1]}}

\begin{document}

\title{Galant: A Graph Algorithm Animation Tool\thanks{
    This work is the result of a senior design project at North Carolina
    State University, Department of Computer Science.
    The first author conceived and guided the project while the others
    provided the implementation details.
    We thank Dr. Robert Fornaro and Margaret Heil, the instructors, for
    coordinating the project(s) and providing instruction in communication
    skills, respectively. Ignacio Dominguez provided valuable technical
    support.
    An alpha testing version of the software is available on request.
    %% provided the recipient (i)~gives a report on how it was used;
    %% and (ii) gives us feedback on features that were desirable, additional
    %% features that would make Galant more useful, bugs, and annoyances.
  }
}
\author{Matthias Stallmann\thanks{Matthias Stallmann, Department of Computer
    Science, North Carolina State University, Raleigh NC, 29695-8206.
    email: \texttt{mfms@ncsu.edu}}
  \and Jason Cockrell
  \and Tynan Devries\thanks{SAS Institute, Cary NC.}
  \and Alexander McCabe\thanks{Elsinore Technologies, Raleigh NC.}
  \and Michael Owoc\thanks{IBM, Durham NC.}
}

\maketitle

\begin{abstract}

\input{abstract}

\end{abstract}

\section{Background}

\input{background}

\section{Related work}

\input{related_work}

\input{galant_features}

%% \section{Research applications}

%% \input{research_applications}

\section{Future work}

\input{future_work}

\bibliographystyle{siam}

\bibliography{Z_GD}

\appendix

\section{Animation examples}
\label{sec:animations}
\input{animation_examples}

\section{User documentation}
\label{sec:user_documentation}
\input{user_documentation}

\section{Programmer guide}
\label{sec:programmer_guide}
\input{programmer_guide}

\section{Known Bugs and Annoyances}
\label{sec:bugs}
\input{bugs_and_annoyances}

\end{document}

% [Last modified: 2013 07 12 at 21:05:53 GMT]
