
\subsubsection*{Input/output}

\begin{itemize}

\item
Weights, if not specified in the input are automatically initialized to 0 instead of, for example, not being displayed at all.

\item
It is impossible to close a tab if it is the only graph or
algorithm -- you must have at least one graph and at least one algorithm open
at any given time.

\item
Tabs for graphs and algorithms are often hard to deal with: (a) if you reread
a graph or algorithm, it appears twice; (b) you can only run an algorithm on a graph if the tabs for the two appear at the top of the window at the same time
-- although the tabs bar can be ``scrolled'',
this may be impossible if there are other intervening graphs and algorithms.

\item
If you close a tab or window you are asked if you want to save changes.
This does not occur, however, if you quit Galant entirely.

\item
When you save the current state of an animation using \texttt{File->Export}
on the graph panel some of the information might not be saved.
In particular, weights and labels are saved, but highlighting is not (since it is not equivalent to an actual color change of a node or edge).

\end{itemize}

\subsubsection*{Graph editing\footnote{
Galant's program editor  is primitive, but
programs can easily be edited externally.}}

\begin{itemize}

\item
When a node is deleted, the id numbers are resequenced so that they run from
0 to $n-1$, where $n$ is the number of nodes.

\item
It is not possible to change the thickness of an edge or node boundary directly
from the editor or an algorithm nor is it possible to change the fill color of a node.
The only way to change these properties is via highlighting, selecting, and
marking nodes/edges during the animation.

\item
A change in color of an edge does not appear to take place while the edge
is selected for editing -- its color stays blue during that time.
Only when focus is no longer on the edge does it change color.

\item
If the user changes the weight of a node or an edge by typing in the text box instead of using the spinner, the weight does not change unless the spinner is
also used (for example, one can enter the desired weight, increment it using the spinner, and then decrement it again).
Sometimes an exception is thrown if the text box is used instead of the spinner. 

\end{itemize}

\subsubsection*{Compilation and execution}

\begin{itemize}

\item
The macro preprocessor is oblivious to comments. Thus, the placement or contents of a comment may cause it to throw an exception.
Such exceptions have been known to occur when comments appear inside functions
or contain constructions that are subject to macro expansion.
For example, sometimes the word \texttt{for} in a comment causes problems.

\item Every once in a while an exception occurs when an error-free algorithm
  is executed or when Galant intitally fires up, but it is possible to step
  through the animation normally after hitting the \textsf{Continue} button.

\item 
  The \texttt{function} construct appears not to work if the arguments or
  return value are arrays (or lists?).
  Macro expansion for this construct is complex.

\item
Compiler error messages can be cryptic (but at least they refer to the correct
line numbers). Because of the macro preprocessing, it may be necessary to look at the console to get an idea of what is causing a particular error.

\item
There is no way for the user to change anything about the display while stepping through the animation, such as, for example, changing visibility of labels and weights,
or positions of nodes to make important features more visible.

\item
There is no way for the animation program to specify that labels and/o weights
should be displayed or hidden.

\item
There is no way to execute the animation in a continuous fashion with a
controllable speed. The current workaround is the use of arrow keys as keyboard
shortcuts for stepping forward or backward -- these can be held down to
generate multiple steps in rapid succession, but the user must click on the
graph panel before starting the animation; for some reason, there is no
reaction to the shortcuts if the first mouse click is on one of the arrows at the bottom of the graph window.

\end{itemize}

