\section{Some existing Galant animations}
\label{app:galant_animations}

The purpose of these examples is mainly to illustrate the capabilities of
Galant. Many other algorithms are easily animated and some, such as
breadth-first search and Kruskal's minimum spanning tree algorithm have
already been implemented.

\subsection{Depth-first search}

Fig.~\ref{fig:dfs} illustrates an animation of depth-first search.
The code, in the current version of Galant, requires a few (four to be exact) lines lines at the beginning to set up the global variable \verb$time$ and the arrays \verb$discovery$ and \verb$finish$.
See the programmer guide (Appendix~\ref{app:programmer_guide}) for more
detailed instructions on how to do this properly.

The animation follows the conventions of Cormen et al. --- tree edges are selected (shown as thick red lines) and non-tree edges are labeled as B = back edge,
F = forward edge, and C = cross edge.
White nodes, not yet visited,
are neither marked nor selected;
gray nodes, visited but the visit is not finished, are selected
(thick red boundary);
and black nodes, visit is completed, are both marked and selected (red boundary and gray fill).
Labels on nodes indicate the discovery and finish times, separated by a slash.

Note that the animation begins with the statement\\
\hspace{3em} \verb+setDirected(true)+\\
which ensures that the graph is displayed and treated as a directed graph.
Depth-first search on an undirected graph requires keeping track of parents of visited nodes
or distinguishes visited from unvisited edges.

\input{Y_dfs}

\subsection{Dijkstra's algorithm}

Fig.~\ref{fig:dijkstra} illustrates an animation of Dijkstra's shortest path algorithm.

\cmt{mention the role of node 0, the fact that decrease key is done in clumsy fashion to avoid a pointer into the queue (the actual decreaseKey function is not shown; not a big deal for small graphs, but
could be avoided if creator provided a binary heap implementation (which could
be an external class) and added an integer attribute.}

\input{Y_dijkstra}



\subsection{Bubble sort}

% [Last modified: 2013 07 12 at 21:05:01 GMT]
