One of the simplest algorithms we have implemented is Dijkstra's algorithm
for the single source shortest paths problem.
Fig.~\ref{fig:dijkstra}
shows most of the implementation of the animation of Dijkstra's algorithm.
At every step the nodes already in the shortest path tree are \emph{visited}
(gray shading) and the nodes that have been encountered (but are not in the tree)
are \emph{selected }(thick red boundary).\footnote{
  Because these node and edge states both guide the logic of the algorithm
  and how the nodes/edges are displayed, the nomenclature has become awkward.
  A better way to handle the situation is to add initial declarations
  that specify color, thickness, and, in case of nodes, fill color, for each logical state.
  A logical state would have a name, e.g., \emph{InTree},
  and be automatically provided with a setter (Boolean argument),
  e.g., \emph{setInTree}, and a logical test, e.g., \emph{isInTree}.
}
Selected \emph{edges} (thick red) represent the current shortest paths to all
encountered nodes;
they are the edges of a shortest path tree when the algorithm is done.
The same algorithm animation
works for both directed and undirected graphs, as illustrated
in Figs.~\ref{fig:dijkstra_directed} and~\ref{fig:dijkstra_undirected}.
The user can toggle between the directed and undirected versions of a graph
via push of the appropriate button.
The functions \emph{beginStep} and \emph{endStep} define the points at which
the exploration of the algorithm stops its forward or backward motion.
In their absence, any state change (mark, select, change in weight, etc.)
constitutes a step, which, in some cases, can force the user
to do excessive ``stepping'' to move past uninteresting state changes.

\input{Y_dijkstras_algorithm}

\input{Y_dijkstra_directed}

\input{Y_dijkstra_undirected}

The macro \emph{for\_outgoing($v,e,w$)}
creates a loop whose body is executed once
for each edge leading out of $v$; in the body, $e$ refers to the current edge
and $w$ to the other endpoint (any other variable names can be used).
In an undirected graph the term \emph{outgoing} applies to all incident
edges.\footnote{
Also provided are \emph{for\_incoming} and \emph{for\_adjacent};
the latter applies to all incident edges, even for directed graphs.
}

The difference between what the algorithm does on a directed versus an undirected graph is evident in the figures.
The edge \emph{from} node~3 to node~2 in the directed graph becomes an
edge \emph{between} the two nodes in the undirected form of the same graph.
Thus, in the undirected version, when node~2 is added to the tree
it also causes the distance from the source, node~0, to node~3 to be updated,
via the path through node~2.
These snapshots come from the executions of the \emph{same algorithm} on the
\emph{same graph}.
The only difference is that the explorer toggled from the directed to
the undirected
interpretation of the graph.
